\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-04-14 eg6-letters/NIM/EG6_RTPC_NIM.tex
Finished reread of NIM

- Mention trigger time.

@@

@@

The parametrization of the drift speed and the drift paths were carried out 
using only the TDCs, gain calibration goal is to equalize the gains in ADC/MeV 
over the full RTPC. The gain of each pad is the ratio between the actual 
deposited energy and the registered ADC value. We implemented three different 
methods to extract such gains, with varying degrees of success.

The first method is based on the measured average $\small{\frac{dE}{dX}}$
and the expected value based on the Bethe-Bloch formula (PDG REF).  The total energy
deposition for each $^4$He track is calculated as the sum of all ADC hits
attributed to the track.  The path length of that energy deposition is
calculated as the distance along the reconstructed trajectory between the first
and last hits on the track and corrected to account for regions
of bad channels along the track.  The measured average energy loss is then the
ratio of deposited energy and path length.  The Bethe-Bloch formula is
evaluated at the track's reconstructed momentum, and the ratio with the measured
$\small{\frac{dE}{dX}}$ is taken as a gain scaling factor.  This average gain is
attributed equally to each readout channel contributing to the track, and each
channel's gain is averaged over many tracks.  This gain calibration method is
inherently iterative, and, while it does result in measured $\small{\frac{dE}{dX}}$
much closer to the expected Bethe-Bloch curve than prior to calibration, the
results were not very satisfactory.

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by simulated tracks (using the same elastic events than for the drift 
paths calibration).  This requires a very good GEANT4 simulation 
including drift paths, but also drift and amplification spreads of the charges 
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
can compare simulation to data on an event by event basis as in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is defined 
as the ratio of the measured ADCs to the simulated deposited energy.

In the third method, the ADC value of each pad is compared to 
the ADCs of the other pads inside each track.
(Detail of 3rd method: Mohammad)

Do you want to show $log(dE/dX)$?

@@

\begin{figure}[tb]
   \includegraphics[scale=0.26]{fig/dedx_p_exp_1st.png}
   \caption{$\small{\frac{dE}{dX}}$ vs. p distribution for the left half of the 
      RTPC (left) and for the right half (right). Here, $\small{\frac{dE}{dX}}$ 
   is calculated using the gains of the first method.  The lines are 
theoretical expectations from Bethe-Bloch formula for $^4$He, $^3$He, $^3$H and 
$^2$H (d).}
\label{fig:dedx_p_exp_1st}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.26]{fig/dedx_p_exp_2nd.png}
\caption{$\small{\frac{dE}{dX}}$ vs. p distribution for the left half of the 
   RTPC (left) and for the right half (right). Here, $\small{\frac{dE}{dX}}$ is 
   calculated using the gains of the second method.  The lines are theoretical 
   expectations from Bethe-Bloch formula for $^4$He, $^3$He, $^3$H and $^2$H 
(d).}
\label{fig:dedx_p_exp_2nd}
\end{figure}

A set of gains has been extracted from each method, in 
figure~\ref{fig:dedx_p_exp_1st}, we have $\small{\frac{dE}{dX}}$ calculated 
using the first method's gains, while in Figure~\ref{fig:dedx_p_exp_2nd} we use 
the second method's gains. From these, we concluded that the gains of the 
second method match best the theoretical lines. This is not surprising since 
calibrating with full tracks leads to mixing the gains from the different pads.  
So it seems that even with several iterations, we could not decipher properly 
the effects of individual pads with this method.

@@

%\begin{figure}[tb]
%\hspace{-0.4cm}
%\includegraphics[scale=0.27]{fig/noisy_pad_after_rejection.png}
%\caption{}
%\label{fig:noisy_pad_after_rejection}
%\end{figure}

@@

In order to reconstruct tracks we first select good hits. This means rejecting 
out-of-time hits and hits linked to the electronic noise. The second step is 
space-reconstructing the hits using the extracted drift speed and drift path 
parameters. For each registered hit, we calculate a position of emission from 
the recorded time (TDC) and the position of the recording pad. The third step 
is to create chains of hits. The maximum distance between two close adjacent 
hits has to be less than 10.5 mm to chain them, this roughly correspond to 
neighbors and next to neighbors. Then, we fit the chains that have a minimum of 
10 hits. We make the fit in two iterations, first, all the hits of the chain 
together with the beam line are fitted with a helix. For the second iteration, 
the hits that are 5 mm or farther from the first fit are excluded.

@@

(We need some performance studies here: Mohammad)

@@

>>>>

- Mention trigger time T=15.

@@

- Mention max rate we ran

@@

The gain of each pad is defined as the ratio between the actual 
deposited energy and the registered ADC value. We tested two different 
methods to extract such gains, with varying degrees of success.

The first method is based on the measured average $\small{\frac{dE}{dX}}$
and the expected value based on the Bethe-Bloch formula. The total energy
deposition for each $^4$He track is calculated as the sum of all ADC hits
attributed to the track. The path length of that energy deposition is
calculated as the distance along the reconstructed trajectory between the first
and last hits on the track and corrected to account for regions
of bad channels along the track. The measured average energy loss is then the
ratio of deposited energy and path length. The Bethe-Bloch formula is
evaluated at the track's reconstructed momentum, and the ratio with the measured
$\small{\frac{dE}{dX}}$ is taken as a gain scaling factor. This average gain is
attributed equally to each readout channel contributing to the track, and each
channel's gain is averaged over many tracks. This gain calibration method is
inherently iterative, and, while it does result in measured $\small{\frac{dE}{dX}}$
much closer to the expected Bethe-Bloch curve than prior to calibration, the
results shown in figure~\ref{fig:dedx_p_exp_1st} are not very satisfactory.

\begin{figure}[tb]
   \includegraphics[scale=0.26]{fig/dedx_p_exp_1st.png}
   \caption{$\small{\frac{dE}{dX}}$ vs. $p$ distribution for the left half of the 
      RTPC (left) and for the right half (right). Here, $\small{\frac{dE}{dX}}$ 
   is calculated using the gains of the first method.  The lines are 
theoretical expectations from Bethe-Bloch formula for $^4$He, $^3$He, $^3$H and 
$^2$H (d).{\bf \color{red} Mohammad: Did you multiply p/c by 2 in the figure? If no 
change the caption, if yes check remove the single charge nuclei from the figure 
because there is a consistency issue.}}
\label{fig:dedx_p_exp_1st}
\end{figure}

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by similar simulated tracks (using the same elastic events than for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is defined 
as the ratio of the measured ADCs to the simulated deposited energy.
Then, the ADC value of each pad is compared to 
the ADCs of the other pads inside each track.
{\bf \color{red} Mohammad: Detail are missing here.}
Results of this second method are shown in Figure~\ref{fig:dedx_p_exp_2nd},
from these we concluded that the gains of the 
second method match best the theoretical lines.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.26]{fig/dedx_p_exp_2nd.png}
\caption{$\small{\frac{dE}{dX}}$ vs. $p$ distribution for the left half of the 
   RTPC (left) and for the right half (right). Here, $\small{\frac{dE}{dX}}$ is 
   calculated using the gains of the second method.  The lines are theoretical 
   expectations from Bethe-Bloch formula for $^4$He, $^3$He, $^3$H and $^2$H 
(d).{\bf \color{red} Mohammad: Did you multiply p/c by 2 in the figure? If no 
change the caption, if yes check remove the single charge nuclei from the figure 
because there is a consistency issue.}}
\label{fig:dedx_p_exp_2nd}
\end{figure}

@@

@@

In order to reconstruct tracks we first select good hits. This means rejecting 
out-of-time hits and hits linked to the electronic noise. The second step is 
to reconstruct the spacial origin of the hits using the extracted drift speed and drift path 
parameters. For each registered hit, we calculate a position of emission from 
the recorded time and the position of the recording pad. The third step 
is to create chains of hits. The maximum distance between two close adjacent 
hits has to be less than 10.5 mm to chain them, this roughly correspond to 
neighbors and next to neighbors. Then, we fit the chains that have a minimum of 
10 hits. We make the fit in two iterations, first, all the hits of the chain 
together with the beam line are fitted with a helix. For the second iteration, 
the hits that are 5 mm or farther from the first fit are excluded.

@@

{\bf \color{red} Mohammad: We need some performance studies here, I remember we had
figures showing efficiency, we need to add them.}

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure helium-4 nuclei in high rate environnement. The operation
of the detector was successful and allowed to detect helium nuclei at a rate
of several kHz in coincidence with the detection of high energy electrons and 
photons in the CLAS spectrometer.
